---
title: "Announcing shiny.webawesome: a web UI package for R/Shiny"
engine: knitr
format:
  html:
    embed-resources: false
    toc: true
    page-layout: full
filters:
  - shinylive
---

```{=html}
<style>
#title-block-header .title {
  font-size: 1.78rem !important;
  line-height: 1.15;
}
</style>
```

::: {.callout-note title="TL;DR"}
`shiny.webawesome` brings Web Awesome to R Shiny through generated wrappers,
reactive bindings, and a bundled runtime. It aims for complete component
support while staying close enough to upstream that the Web Awesome docs and
examples are directly useful in everyday package use.

<a href="https://CRAN.R-project.org/package=shiny.webawesome" target="_blank" rel="noopener noreferrer">CRAN</a> | <a href="https://mbanand.r-universe.dev/shiny.webawesome" target="_blank" rel="noopener noreferrer">R-universe</a> | <a href="https://www.shiny-webawesome.org" target="_blank" rel="noopener noreferrer">Package website</a> | <a href="https://github.com/mbanand/shiny.webawesome" target="_blank" rel="noopener noreferrer">Source repository</a>
:::

## Background

`shiny.webawesome` started from a perceived gap: Shiny would benefit from a
UI library that feels modern, visually polished, and broad enough to support
a full app coherently. Web Awesome was a strong fit because it combines rich
components, layout and styling utilities, and detailed upstream documentation
with a standards-based web-components structure that is straightforward to
track from R. That makes it easier for the package to stay close to upstream
while still fitting naturally into Shiny.

## The Whole Game

Here's a simple, complete example app, rendered via Shinylive.

```{=html}
<style>
#quarto-document-content div.sourceCode,
#quarto-document-content pre.sourceCode {
  max-height: 30rem;
  overflow: auto;
  border-radius: 16px;
}

.example-code-viewer {
  margin: 1rem 0 2rem;
  border: 1px solid #d7deea;
  border-radius: 16px;
  overflow: hidden;
  background: #fbfcfe;
  box-shadow: 0 16px 40px rgba(15, 23, 42, 0.06);
}

.example-code-viewer-header {
  display: grid;
  grid-template-columns: 1fr auto;
  align-items: stretch;
  background: linear-gradient(180deg, #f7f9fc, #eef3f9);
  border-bottom: 1px solid #d7deea;
}

.example-code-tabs {
  display: flex;
  flex: 1 1 auto;
}

.example-code-tab {
  appearance: none;
  border: 0;
  background: transparent;
  flex: 1 1 0;
  min-width: 4.5rem;
  padding: 1rem 1.25rem;
  font-weight: 700;
  font-size: 0.98rem;
  text-align: center;
  color: #526077;
  border-right: 1px solid #d7deea;
}

.example-code-tab.is-active {
  background: #ffffff;
  color: #0f172a;
}

.example-code-copy {
  appearance: none;
  border: 0;
  min-width: 3.25rem;
  border-left: 1px solid #d7deea;
  background: transparent;
  color: #526077;
  padding: 0 1rem;
  position: relative;
}

.example-code-copy.is-copied::after,
.example-code-copy.is-error::after {
  position: absolute;
  top: calc(100% + 0.4rem);
  right: 0.5rem;
  padding: 0.28rem 0.5rem;
  border-radius: 999px;
  background: #0f172a;
  color: #ffffff;
  font-size: 0.78rem;
  font-weight: 700;
  line-height: 1;
  white-space: nowrap;
  z-index: 2;
}

.example-code-copy.is-copied::after {
  content: "Copied";
}

.example-code-copy.is-error::after {
  content: "Copy failed";
}

.example-code-copy:hover,
.example-code-tab:hover {
  background: rgba(255, 255, 255, 0.7);
}

.example-code-pane {
  display: none;
}

.example-code-pane.is-active {
  display: block;
}

.example-code-pre {
  margin: 0;
  height: 400px;
  overflow: auto;
  padding: 1rem 1.1rem;
  background: #ffffff;
  color: #0f172a;
  font-size: 0.92rem;
  line-height: 1.55;
  font-family: "SFMono-Regular", "Consolas", "Liberation Mono", monospace;
}
</style>

<script>
document.addEventListener("DOMContentLoaded", () => {
  document.querySelectorAll(".example-code-viewer").forEach((viewer) => {
    const tabs = Array.from(viewer.querySelectorAll(".example-code-tab"));
    const panes = Array.from(viewer.querySelectorAll(".example-code-pane"));
    const copyButton = viewer.querySelector(".example-code-copy");
    let feedbackTimeout = null;

    function showCopyFeedback(state) {
      if (!copyButton) return;

      copyButton.classList.remove("is-copied", "is-error");
      if (feedbackTimeout) {
        window.clearTimeout(feedbackTimeout);
      }

      copyButton.classList.add(state);
      feedbackTimeout = window.setTimeout(() => {
        copyButton.classList.remove("is-copied", "is-error");
      }, 1200);
    }

    function activatePane(target) {
      tabs.forEach((tab) => {
        tab.classList.toggle("is-active", tab.dataset.target === target);
      });
      panes.forEach((pane) => {
        pane.classList.toggle("is-active", pane.dataset.pane === target);
      });
    }

    tabs.forEach((tab) => {
      tab.addEventListener("click", () => activatePane(tab.dataset.target));
    });

    if (copyButton) {
      copyButton.addEventListener("click", async () => {
        const activePane = viewer.querySelector(".example-code-pane.is-active code");
        if (!activePane) return;

        try {
          await navigator.clipboard.writeText(activePane.textContent);
          showCopyFeedback("is-copied");
        } catch (error) {
          showCopyFeedback("is-error");
        }
      });
    }
  });
});
</script>
```

```{r}
#| echo: false
#| results: asis
article_dir <- dirname(knitr::current_input(dir = TRUE))
app_r_lines <- readLines(file.path(article_dir, "app.R"), warn = FALSE)
app_css <- paste(readLines(file.path(article_dir, "app.css"), warn = FALSE), collapse = "\n")
app_js <- paste(readLines(file.path(article_dir, "app.js"), warn = FALSE), collapse = "\n")

replacement_lines <- c(
  "show_shinylive_badge <- TRUE",
  sprintf('css_text <- %s', encodeString(app_css, quote = '"')),
  sprintf('js_text <- %s', encodeString(app_js, quote = '"'))
)

start_idx <- grep("^if \\(!exists\\(\"css_text\"", app_r_lines)
end_idx <- grep("^}$", app_r_lines)
if (length(start_idx) != 1L || length(end_idx) == 0L) {
  stop("Could not locate css/js bootstrap block in announcement/app.R")
}
end_idx <- end_idx[end_idx > start_idx][1]
app_r_lines <- c(app_r_lines[seq_len(start_idx - 1)], replacement_lines, app_r_lines[(end_idx + 1):length(app_r_lines)])

cat("```{shinylive-r}\n")
cat("#| standalone: true\n")
cat("#| viewerHeight: 810\n\n")
cat(paste(app_r_lines, collapse = "\n"))
cat("\n```\n")
```

<!--
The generated shinylive-r block above inlines app.css and app.js as literal
strings so the live app works in Shinylive/webR, while the sibling files remain
available as copyable assets for a standalone local app.
-->

The code for the above example is shown below.

This example showcases many of the facilities available in the package:

- a visually rich component library
- direct use of Web Awesome layout utilities such as `wa-stack`,
  `wa-cluster`, `wa-gap-*`, and `wa-align-*` classes
- styling through Web Awesome design tokens and classes such as
  `--wa-color-*`, `--wa-font-*`, and `wa-body-*`
- reactive Shiny input bindings
- helpers for executing commands on HTML elements, setting properties, and
  injecting simple JavaScript snippets

```{r}
#| echo: false
#| results: asis
article_dir <- dirname(knitr::current_input(dir = TRUE))
app_r <- paste(readLines(file.path(article_dir, "app.R"), warn = FALSE), collapse = "\n")
app_css <- paste(readLines(file.path(article_dir, "app.css"), warn = FALSE), collapse = "\n")
app_js <- paste(readLines(file.path(article_dir, "app.js"), warn = FALSE), collapse = "\n")

escape_html <- function(x) {
  htmltools::htmlEscape(x)
}

cat(
  '<div class="example-code-viewer">',
  '<div class="example-code-viewer-header">',
  '<div class="example-code-tabs">',
  '<button class="example-code-tab is-active" type="button" data-target="r">R</button>',
  '<button class="example-code-tab" type="button" data-target="css">CSS</button>',
  '<button class="example-code-tab" type="button" data-target="js">JS</button>',
  '</div>',
  '<button class="example-code-copy" type="button" aria-label="Copy active code"><i class="bi bi-copy"></i></button>',
  '</div>',
  '<div class="example-code-pane is-active" data-pane="r"><pre class="example-code-pre"><code>',
  escape_html(app_r),
  '</code></pre></div>',
  '<div class="example-code-pane" data-pane="css"><pre class="example-code-pre"><code>',
  escape_html(app_css),
  '</code></pre></div>',
  '<div class="example-code-pane" data-pane="js"><pre class="example-code-pre"><code>',
  escape_html(app_js),
  '</code></pre></div>',
  '</div>',
  sep = ""
)
```

## Design Philosophy

`shiny.webawesome` is designed to stay close to upstream Web Awesome. Most
component wrappers are generated from Web Awesome metadata, which helps
preserve upstream names, structure, and behavior while translating the
interface into normal R conventions such as `snake_case`.

That close alignment has a practical benefit: when you want deeper details,
examples, or component-specific guidance, you can usually go straight to the
upstream Web Awesome documentation and apply what you find directly in
`shiny.webawesome`. The package currently supports all Web Awesome components,
so the upstream docs are a practical reference for day-to-day use.

To support the server-client model of Shiny, the package adds a small set of
page and layout helpers, curated reactive bindings, and a narrow command layer
for cases where browser-side interaction goes beyond the generated wrappers.

The result is a package with a clear default path. Use generated wrappers for
ordinary UI, use bindings for meaningful reactive state, and reach for
commands or small JavaScript glue when the app needs them.

## Shiny Bindings

`shiny.webawesome` does not forward every browser event and every detail of
component telemetry into Shiny. Much component state and interaction detail is
better handled locally in the browser rather than turned into server messages.
Consequently, the package exposes only a curated set of Shiny bindings that
fit Shiny's reactive model, with an emphasis on meaningful committed state
rather than low-level browser event streams.

In the most common case, a binding publishes a durable semantic value. A
select reports its current value, a dialog can report whether it is open, and
a tree can report the currently selected item ids. The key idea is that Shiny
receives the state the app actually cares about, not the raw event name that
happened to produce it.

Some components are better treated as actions than values. A button is the
clearest example: in Shiny, it behaves like a Shiny action input, with each
click producing a new input event. A small number of components need both
action semantics and a separate value. A dropdown, for example, may need to
trigger reactivity on every choice, including repeated selections of the same
item, while also exposing the latest selected value.

This design keeps reactive messaging to the server smaller, clearer, and
easier to reason about. If an interaction belongs naturally in Shiny's input
model, `shiny.webawesome` will expose it as a binding. If it is more naturally
a browser-side concern, it is usually a better fit for the command layer or a
small amount of JavaScript glue.

For the full binding categories, semantics, and examples, see the package
article: [Shiny Bindings](https://www.shiny-webawesome.org/articles/shiny-bindings.html).

## Command API

`shiny.webawesome` covers the most common interaction patterns through
generated wrappers, Shiny bindings, and update helpers. But sometimes an app
still needs to reach into a live browser element directly: set a property,
call a method, or add a small browser-local JavaScript snippet.

For those cases, the package provides a narrow command API. The two main
server-side helpers are `wa_set_property()` and `wa_call_method()`. They let
Shiny code send one-way commands to a browser element identified by `id`,
either by assigning a value to a live property or invoking a browser-side
method.

If a component already has a binding or update helper, that should usually
remain the first choice. The command layer is for the cases that fall just
outside those built-in paths, where the simplest solution is still to tell the
existing browser component to do one specific thing.

The package also includes `wa_js()` for a different kind of job: small,
app-local JavaScript glue. That is useful when the missing piece is
browser-side logic such as listening for an event, reading live component
state, or publishing a derived value back to Shiny with
`Shiny.setInputValue()`.

For more detail and examples, see the package article: [Command API](https://www.shiny-webawesome.org/articles/command-api.html).

## Conclusion

`shiny.webawesome` brings a visually rich component library into Shiny while
staying close to upstream Web Awesome. That combination gives polished
components, useful layout and styling utilities, and a workflow where
upstream documentation and examples remain directly relevant throughout app
development.

For more examples, longer articles, and full reference material, see the
package website: [shiny-webawesome.org](https://www.shiny-webawesome.org).
