---
title: "shiny.webawesome announcement prototype"
engine: knitr
format:
  html:
    embed-resources: false
    toc: true
    page-layout: full
filters:
  - shinylive
---

::: {.callout-note}
This is a working prototype for the eventual r-posts.com / R-bloggers
announcement. The first goal is to prove that Quarto, Shinylive, webR,
R-universe, and the `shiny.webawesome` runtime can all work together.
:::

## TL;DR

`shiny.webawesome` brings Web Awesome components to R Shiny through generated
R wrappers, Shiny bindings, and a bundled Web Awesome runtime.

Links to add after CRAN publication:

- CRAN
- Package website
- Source repository
- R-universe

## Background

Placeholder.

## The Whole Game

This example is intentionally tighter: a small data workbench that keeps the
controls visible, keeps the preview visible, and makes every interaction show
up immediately in the active tab.

```{=html}
<style>
#quarto-document-content div.sourceCode,
#quarto-document-content pre.sourceCode {
  max-height: 30rem;
  overflow: auto;
  border-radius: 16px;
}

.example-code-viewer {
  margin: 1rem 0 2rem;
  border: 1px solid #d7deea;
  border-radius: 16px;
  overflow: hidden;
  background: #fbfcfe;
  box-shadow: 0 16px 40px rgba(15, 23, 42, 0.06);
}

.example-code-viewer-header {
  display: grid;
  grid-template-columns: 1fr auto;
  align-items: stretch;
  background: linear-gradient(180deg, #f7f9fc, #eef3f9);
  border-bottom: 1px solid #d7deea;
}

.example-code-tabs {
  display: flex;
}

.example-code-tab {
  appearance: none;
  border: 0;
  background: transparent;
  padding: 0.9rem 1.1rem;
  font-weight: 700;
  color: #526077;
  border-right: 1px solid #d7deea;
}

.example-code-tab.is-active {
  background: #ffffff;
  color: #0f172a;
}

.example-code-copy {
  appearance: none;
  border: 0;
  border-left: 1px solid #d7deea;
  background: transparent;
  color: #526077;
  padding: 0 1rem;
  min-width: 3.25rem;
}

.example-code-copy:hover,
.example-code-tab:hover {
  background: rgba(255, 255, 255, 0.7);
}

.example-code-pane {
  display: none;
}

.example-code-pane.is-active {
  display: block;
}

.example-code-pre {
  margin: 0;
  height: 400px;
  overflow: auto;
  padding: 1rem 1.1rem;
  background: #ffffff;
  color: #0f172a;
  font-size: 0.92rem;
  line-height: 1.55;
  font-family: "SFMono-Regular", "Consolas", "Liberation Mono", monospace;
}
</style>

<script>
document.addEventListener("DOMContentLoaded", () => {
  document.querySelectorAll(".example-code-viewer").forEach((viewer) => {
    const tabs = Array.from(viewer.querySelectorAll(".example-code-tab"));
    const panes = Array.from(viewer.querySelectorAll(".example-code-pane"));
    const copyButton = viewer.querySelector(".example-code-copy");

    function activatePane(target) {
      tabs.forEach((tab) => {
        tab.classList.toggle("is-active", tab.dataset.target === target);
      });
      panes.forEach((pane) => {
        pane.classList.toggle("is-active", pane.dataset.pane === target);
      });
    }

    tabs.forEach((tab) => {
      tab.addEventListener("click", () => activatePane(tab.dataset.target));
    });

    if (copyButton) {
      copyButton.addEventListener("click", async () => {
        const activePane = viewer.querySelector(".example-code-pane.is-active code");
        if (!activePane) return;

        await navigator.clipboard.writeText(activePane.textContent);
      });
    }
  });
});
</script>
```

```{r}
#| echo: false
#| results: asis
article_dir <- dirname(knitr::current_input(dir = TRUE))
app_r_lines <- readLines(file.path(article_dir, "app.R"), warn = FALSE)
app_css <- paste(readLines(file.path(article_dir, "app.css"), warn = FALSE), collapse = "\n")
app_js <- paste(readLines(file.path(article_dir, "app.js"), warn = FALSE), collapse = "\n")

replacement_lines <- c(
  "show_shinylive_badge <- TRUE",
  sprintf('css_text <- %s', encodeString(app_css, quote = '"')),
  sprintf('js_text <- %s', encodeString(app_js, quote = '"'))
)

start_idx <- grep("^if \\(!exists\\(\"css_text\"", app_r_lines)
end_idx <- grep("^}$", app_r_lines)
if (length(start_idx) != 1L || length(end_idx) == 0L) {
  stop("Could not locate css/js bootstrap block in announcement/app.R")
}
end_idx <- end_idx[end_idx > start_idx][1]
app_r_lines <- c(app_r_lines[seq_len(start_idx - 1)], replacement_lines, app_r_lines[(end_idx + 1):length(app_r_lines)])

cat("```{shinylive-r}\n")
cat("#| standalone: true\n")
cat("#| viewerHeight: 810\n\n")
cat(paste(app_r_lines, collapse = "\n"))
cat("\n```\n")
```

<!--
The generated shinylive-r block above inlines app.css and app.js as literal
strings so the live app works in Shinylive/webR, while the sibling files remain
available as copyable assets for a standalone local app.
-->

The code for the above example is shown below.

```{r}
#| echo: false
#| results: asis
article_dir <- dirname(knitr::current_input(dir = TRUE))
app_r <- paste(readLines(file.path(article_dir, "app.R"), warn = FALSE), collapse = "\n")
app_css <- paste(readLines(file.path(article_dir, "app.css"), warn = FALSE), collapse = "\n")
app_js <- paste(readLines(file.path(article_dir, "app.js"), warn = FALSE), collapse = "\n")

escape_html <- function(x) {
  htmltools::htmlEscape(x)
}

cat(
  '<div class="example-code-viewer">',
  '<div class="example-code-viewer-header">',
  '<div class="example-code-tabs">',
  '<button class="example-code-tab is-active" type="button" data-target="r">R</button>',
  '<button class="example-code-tab" type="button" data-target="css">CSS</button>',
  '<button class="example-code-tab" type="button" data-target="js">JS</button>',
  '</div>',
  '<button class="example-code-copy" type="button" aria-label="Copy active code"><i class="bi bi-copy"></i></button>',
  '</div>',
  '<div class="example-code-pane is-active" data-pane="r"><pre class="example-code-pre"><code>',
  escape_html(app_r),
  '</code></pre></div>',
  '<div class="example-code-pane" data-pane="css"><pre class="example-code-pre"><code>',
  escape_html(app_css),
  '</code></pre></div>',
  '<div class="example-code-pane" data-pane="js"><pre class="example-code-pre"><code>',
  escape_html(app_js),
  '</code></pre></div>',
  '</div>',
  sep = ""
)
```

## Design Philosophy

Placeholder.

## Shiny Bindings

Placeholder.

## Command API

Placeholder.

## Conclusion

Placeholder.
